Improving the Image Acquisition Rate of an Atomic Force Microscope Through Compressive Sampling

### 2019-06-06 latex/afm-cs-main.tex
Update writing of control section.

% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is different. For the $X$-axis, we use a
state space controller. It is described fully in \cite{braker_afmmpc_2019} and
achieves a bandwidth of about 150 Hz, which is over 1/3 of the first $X$-axis
resonance at 350 Hz.
\todo{Not sure how to give a deeper summary without getting to much into the weeds.}

The control law used for the $Y$-axis is a simple PI controller of the
form
\begin{equation}
  \frac{U(z)}{E_Y(z)}=D(z) = \frac{K_Iz}{z-1},
  \label{eqn:dzI}
\end{equation}
where, the error signal $E_Y(z) = Y(z)-R_Y$.

\todo{Should I implement something better for $Y$-axis, e.g., cancel complex
  pole-zeros at 350 hz and add notch filter???}

The $Z$-axis control is slightly more complicated than the simple $Y$-axis
integral controller and is described in the following subsection.
\subsection{Z-axis Control Design}\label{sec:zaxis_cs_control}
The main challenge to increasing the $Z$-axis bandwidth is the bending mode
resonance at 215 Hz, which can be seen in the FRF from $u_Z$ to $Z_{d}$ in
Fig.~\ref{fig:z_control} (blue curve). We address this by inverting the complex
pole-zero pair at 215 Hz. Thus, the entire $Z$-axis controller consists of a PI
controller, $D_I(z)$ cascaded with the resonance inversion, $D^{-1}(z)$, as shown in Fig.~\ref{fig:afm_bd_dinv}. The DC-gain
of $D$ is (approximately) the same as the DC-gain of the plant.
The inclusion of $D^{-1}$ allows us to substantially increase the PI gain and
achieve a closed-loop bandwidth of about 480 Hz. This is the black curve in Fig.~\ref{fig:z_control}.

There are multiple advantages to this method: first,
an inverse compensator still makes sense as a feedforward, open-loop compensator
if the feedback path is broken. This situation can and does occur during parts
of the CS cycle. Second, the inverse compensator
requires no tuning as it is defined only by a fit to the measured FRF.

One challenge to the inversion approach is both the system gain and resonance at 215~Hz
changes not only day-to-day but also (and especially) when changing cantilevers and our inversion is not robust to this. However, because $D^{-1}$ is defined purely by a second order fit to a small band of the FRF, updating the compensator can be automated.
\todo{What else to say here?}

@@

\subsection{Feedforward Control}\label{sec:ff_control}
In our AFM, there is a strong coupling between the horizontal axes and vertical deflection. Fig.~\ref{fig:mimo_frf_uxuy} shows a MIMO FRF for $u_X$ and $u_Y$ to the outputs in each axis. The dotted black curves are the open loop responses and the dashed red curves are the nominal closed-loop responses. Though the $Z$-direction inputs are not shown, the closed loop responses include the $Z$-axis controller. Note that, especially for $H_{z,u_X}$, the closed-loop response is barely improved over the open loop response and in particular, there is a sharp resonance at 215~Hz and 505~Hz. To deal with this, we employ two different strategies for CS and raster scanning.

For CS scanning, we insert a feed-forward filter with two notches, one at 215~Hz and the other at 505~Hz. We design a similar feed-forward filter for the $Y$-axis. Both filters are shown in Fig.~\ref{fig:mimo_frf_uxuy} as the solid pink curves and the resulting overall FRF is shown as the solid blue curves. Particularly for $H_{z,u_X}$, these filters result in a dramatic reduction in the cross coupling. A similar affect could be achieve by using a simple low-pass filter. However, to achieve the same 40~dB of attenuation of the mode at 215 Hz, would require, e.g., that a two-pole LPF have its cut-off frequency at about 20~Hz.

For raster scanning, we use a truncated Fourier series, where each Fourier coefficient is scaled by the inverse of the closed-loop FRF at the corresponding frequency. We truncate this series such that the highest frequency is smaller than 200 Hz. See, e.g., \cite{clayton_review_2009} for more details on this strategy. Note that these scheme implies that the effective bandwidth of the $X$-axis controller is higher while raster scanning than while CS-scanning. On the other hand, it also implies that the for faster scan rates, fewer Fourier coefficients are used, and so the approximation to a triangle wave is worse.

>>>>

% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is similar and essentially consists of a set of filters in series with a PI controller. The open-loop frequency response for the $Z$-axis is shown in Fig.~\ref{fig:z_control} (red curve) and in Fig.~\ref{fig:mimo_frf_uxuy} for the $X$ and $Y$ axes (black curves). The first resonance of each axis is the bending mode. For the $Z$ axis, this resonance is at 215~Hz and at approximately 350~Hz for the $X$ and $Y$ axes.

The filter for the $Z$-axis controller inverts the complex-conjugate pole-zero pair at 215~Hz. This allows the gain of the PI controller to be increased substantially. The closed-loop achieves a bandwidth of about 450~Hz, as shown by the black curve in Fig.~\ref{fig:z_control}.

We take a similar approach to the $X$ and $Y$ axes. The filter inverts the bending modes at 350~Hz and adds a notch filter for the main resonant peak at 650~Hz. The resulting closed-loops are shown in the (1,1) and (2,2) positions of Fig.~\ref{fig:mimo_frf_uxuy}. In general, the $X$-axis positioning requirements are more stringent for both CS and raster scanning

Unfortunately, as Fig.~\ref{fig:mimo_frf_uxuy} makes clear, there is a strong cross-coupling between the axes. The coupling from the horizontal plane to the $Z$ axis is particularly problematic and our SISO control scheme does little to attenuate the resonances of $G_{Z,u_X}$ and $G_{Z,u_Y}$. In particular, there are modes in both these transfer functions at 215 and 505 Hz that do not appear in a SISO model of $G_{X,u_x}$ or $G_{y,u_Y}$.

For raster scans, the frequency content of the $X$-axis reference (which is a triangle wave) decays like $1/n^2$ with a gain proportional to the raster frequency. In contrast, the $X$-axis reference for a $\mu$-path scan is composed of a sequence of steps and ramps. The inclusion of the (discontinuous) steps means we should expect the $\mu$-path $X$ reference to decay like $1/n$. 
Thus,  while the cross-coupling can induce $Z$-axis vibrations for faster raster scans, it is especially problematic for $\mu$-path scanning. In principle, one could mitigate these issues with a full MIMO control design. In this work, we opt for a much simpler feedforward scheme described in the next subsection.

% The control law used for the $Y$-axis is a simple PI controller of the
% form
% \begin{equation}
%   \frac{U(z)}{E_Y(z)}=D(z) = \frac{K_Iz}{z-1},
%   \label{eqn:dzI}
% \end{equation}
% where, the error signal $E_Y(z) = Y(z)-R_Y$.
% 
% \todo{Should I implement something better for $Y$-axis, e.g., cancel complex
%   pole-zeros at 350 hz and add notch filter???}

% The $Z$-axis control is slightly more complicated than the simple $Y$-axis
% integral controller and is described in the following subsection.
% \subsection{Z-axis Control Design}\label{sec:zaxis_cs_control}
% The main challenge to increasing the $Z$-axis bandwidth is the bending mode
% resonance at 215 Hz, which can be seen in the FRF from $u_Z$ to $Z_{d}$ in
% Fig.~\ref{fig:z_control} (blue curve). We address this by inverting the complex
% pole-zero pair at 215 Hz. Thus, the entire $Z$-axis controller consists of a PI
% controller, $D_I(z)$ cascaded with the resonance inversion, $D^{-1}(z)$, as shown in Fig.~\ref{fig:afm_bd_dinv}. The DC-gain
% of $D$ is (approximately) the same as the DC-gain of the plant.
% The inclusion of $D^{-1}$ allows us to substantially increase the PI gain and
% achieve a closed-loop bandwidth of about 480 Hz. This is the black curve in Fig.~\ref{fig:z_control}.

% There are multiple advantages to this method: first,
% an inverse compensator still makes sense as a feedforward, open-loop compensator
% if the feedback path is broken. This situation can and does occur during parts
% of the CS cycle. Second, the inverse compensator
% requires no tuning as it is defined only by a fit to the measured FRF.

% One challenge to the inversion approach is both the system gain and resonance at 215~Hz
% changes not only day-to-day but also (and especially) when changing cantilevers and our inversion is not robust to this. However, because $D^{-1}$ is defined purely by a second order fit to a small band of the FRF, updating the compensator can be automated.
% \todo{What else to say here?}

@@

\subsection{Feedforward Control}\label{sec:ff_control}
% In our AFM, there is a strong coupling between the horizontal axes and vertical deflection. Fig.~\ref{fig:mimo_frf_uxuy} shows a MIMO FRF for $u_X$ and $u_Y$ to the outputs in each axis. The dotted black curves are the open loop responses and the dashed red curves are the nominal closed-loop responses. Though the $Z$-direction inputs are not shown, the closed loop responses include the $Z$-axis controller. Note that, especially for $H_{z,u_X}$, the closed-loop response is barely improved over the open loop response and in particular, there is a sharp resonance at 215~Hz and 505~Hz. To deal with this, we employ two different strategies for CS and raster scanning.

For CS scanning, we insert a feed-forward filter with three notches, one at 215~Hz, one at 350~Hz and one at 505~Hz. We design a similar feed-forward filter for the $Y$-axis. Both filters are shown in Fig.~\ref{fig:mimo_frf_uxuy} as the solid pink curves and the resulting overall FRF is shown as the solid blue curves, which show a dramatic reduction in the cross coupling. A similar affect could be achieve by using a simple low-pass filter. However, to achieve the same 40~dB of attenuation of the mode at 215 Hz, would require, e.g., that a two-pole LPF have its cut-off frequency at about 20~Hz.

For raster scanning, we use a truncated Fourier series, where each Fourier coefficient is scaled by the inverse of the closed-loop FRF at the corresponding frequency. We truncate this series such that the highest frequency is smaller than 200 Hz. See, e.g., \cite{clayton_review_2009} for more details on this strategy. Note that this scheme implies that the effective bandwidth of the $X$-axis controller is higher while raster scanning than while CS-scanning. On the other hand, it also implies that the for faster scan rates, fewer Fourier coefficients are used, and so the approximation to a triangle wave is worse.
